---

title: "Cat Postcard Lab: A Reproducible Holiday Incident"
author: "Cat Ops Team"
date: "`r Sys.Date()`"
output:
html_document:
toc: true
toc_depth: 2
df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
echo = TRUE,
message = FALSE,
warning = FALSE
)
set.seed(25) # for reproducible chaos
```

# 🧪 What is this?

This is a **companion R Markdown** file for the *Christmas Postcard With Your Cat* README.

It runs a fully reproducible experiment to answer the world’s most urgent question:

> **How many treats does it take to get one (1) usable photo?**

---

# 🐱 Meet the model

We will simulate a cat photo session using a scientifically rigorous system called **Vibes™**.

```{r}
cat_profile <- list(
name = sample(c("Miso", "Pickles", "Sir Floof", "Chairman Meow", "Bean"), 1),
mood = sample(c("cooperative", "suspicious", "chaotic neutral", "sleepy", "feral holiday gremlin"), 1),
floof_level = sample(1:10, 1),
zoomies_probability = runif(1, 0.1, 0.9)
)
cat_profile



```
Inline report: **`r cat_profile$name`** is currently feeling **`r cat_profile$mood`** with a floof level of **`r cat_profile$floof_level`/10**.
---

# 🎅 Generate a postcard caption

We will generate captions using a carefully curated phrase bank.

```{r}
caption_bank <- list(
opener = c("Meowy Christmas!", "Happy Pawlidays!", "Season’s Greetings!", "Merry Catmas!"),
middle = c(
"May your treats be plentiful",
"May your boxes be infinite",
"May your ornaments survive",
"May your naps be uninterrupted"
),
closer = c("Love," , "With judgement," , "Respectfully," , "PSPSPS," )
)

make_caption <- function(name) {
paste(
sample(caption_bank$opener, 1),
sample(caption_bank$middle, 1),
"\n\n",
sample(caption_bank$closer, 1),
name
)
}

caption <- make_caption(cat_profile$name)
cat(caption)


```

# 🖨️ Print a "postcard" in the console

Because nothing says "holiday" like rendering your cat as text.

```{r}
cat_ascii <- c(
"      /\\_/\\",
"     ( o.o )   * jingles ominously *",
"      > ^ <",
"",
"  [ Christmas Postcard Prototype ]"
)

cat(paste(cat_ascii, collapse = "\n"), "\n\n")
cat(caption, "\n")

```

---

# 💾 Optional: Export postcard text

This chunk writes a small text file you can include in your repo or attach to an email.

```{r}
postcard_text <- paste(cat_ascii, collapse = "\n")
postcard_text <- paste(postcard_text, "\n\n", caption)

outfile <- "postcard.txt"
writeLines(postcard_text, con = outfile)
outfile
```



---

# 📸 Simulate a photo session

We’ll model each photo attempt as a random event influenced by:

* treats deployed
* floof level
* mood (vibes)
* zoomies probability

A photo is considered **usable** if:

* eyes are open
* cat is in frame
* accessories did not start a diplomatic incident

# Convert mood to a "compliance" multiplier

```{r}
# Convert mood to a "compliance" multiplier
mood_multiplier <- function(mood) {
  switch(
    mood,
    cooperative = 1.25,
    sleepy = 1.05,
    suspicious = 0.85,
    "chaotic neutral" = 0.75,
    "feral holiday gremlin" = 0.55,
    0.8
  )
}


simulate_attempt <- function(treats, floof, zoomies_p, mood) {
  vibe <- mood_multiplier(mood)
  
  
  # probabilities (bounded in [0,1])
  in_frame_p <- pmin(0.95, 0.35 + 0.05 * treats + 0.02 * floof) * vibe
  eyes_open_p <- pmin(0.98, 0.55 + 0.03 * treats) * vibe
  no_incident_p <- pmax(0.05, 0.90 - 0.12 * zoomies_p) * (2 - vibe) # chaos increases incidents
  
  
  in_frame <- runif(1) < in_frame_p
  eyes_open <- runif(1) < eyes_open_p
  no_incident <- runif(1) < no_incident_p
  
  
  usable <- in_frame && eyes_open && no_incident
  
  
  outcome <- if (usable) {
    "usable"
  } else if (!in_frame) {
    "cat has left the scene"
  } else if (!eyes_open) {
    "blink captured (art)"
  } else {
    "accessory incident"
  }
  
  
  list(
    usable = usable,
    in_frame = in_frame,
    eyes_open = eyes_open,
    no_incident = no_incident,
    outcome = outcome
  )
}


run_session <- function(n_attempts = 30, starting_treats = 0) {
  treats <- starting_treats
  attempts <- vector("list", n_attempts)
  
  
  for (i in seq_len(n_attempts)) {
    # deploy treats strategically: increase when failure piles up
    if (i %% 5 == 0) treats <- treats + 1
    
    
    attempts[[i]] <- c(list(
      attempt = i,
      treats = treats
    ), simulate_attempt(
      treats = treats,
      floof = cat_profile$floof_level,
      zoomies_p = cat_profile$zoomies_probability,
      mood = cat_profile$mood
    ))
    
    
    # stop early if we got a usable photo
    if (attempts[[i]]$usable) break
  }
  
  
  df <- do.call(rbind, lapply(attempts, function(x) {
    if (is.null(x)) return(NULL)
    data.frame(
      attempt = x$attempt,
      treats = x$treats,
      usable = x$usable,
      outcome = x$outcome,
      stringsAsFactors = FALSE
    )
  }))
  
  
  df
}


session <- run_session(n_attempts = 40)
head(session, 10)
```



## Session verdict

```{r}
usable_row <- which(session$usable)
if (length(usable_row) > 0) {
  i <- usable_row[1]
  paste(
    "SUCCESS after", session$attempt[i], "attempt(s) using",
    session$treats[i], "treat tier(s)."
  )
} else {
  "No usable photos. The cat has won."
}

```

---

# 📊 Plot the chaos

This plot shows the outcomes per attempt.

```{r}
# Encode outcomes for plotting

outcome_levels <- c("usable", "blink captured (art)", "accessory incident", "cat has left the scene")

session$outcome <- factor(session$outcome, levels = outcome_levels)

plot(
  x = session$attempt,
  y = as.numeric(session$outcome),
  xlab = "Attempt",
  ylab = "Outcome (scientific units)",
  yaxt = "n",
  pch = 16
)
axis(2, at = seq_along(outcome_levels), labels = outcome_levels, las = 2)

abline(v = which(session$usable), lty = 2)

```

---

# ✅ Final notes

If this knitted successfully, you have:

* a reproducible simulation
* a caption generator
* a chaos plot
* a freshly exported `postcard.txt`

Most importantly, you have evidence that **science can be festive**.
